---
title: "R Notebook"
output:
  html_document:
    df_print: paged
editor_options:
  markdown:
    wrap: sentence
---

# Import Library

```{r}
library(tidyverse)
library(janitor)
library(readxl)
library(here)
```

\#Import Data

```{r}
clean_data <- read_csv(here("clean_data/full_data.csv"))
```

# Analysis questions

For each question where we ask for bird names, give the bird's common name, scientific name and species abbreviation.

## Question One

Which bird had the most individual sightings?

```{r}
Birds_by_species <- clean_data %>% 
  group_by(species, scientific_name, abbreviation) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))
```

The bird with the most sightings is the Wandering albatross sensu lato PL2 (Diomedea antipodensis / exulans PL2, DIOANTEXU PL2) There were 3104 sightings.

## Question Two

Which bird had the highest total count?

```{r}
clean_data %>% 
  group_by(species, scientific_name, abbreviation) %>% 
  summarise(COUNT = sum(COUNT)) %>% 
  arrange(desc(COUNT))
```

The bird with the highest total count is the black-billed gull (larus bulleri, larbul) with 7732 birds counted

## Question Three

Which bird had the highest total count above a latitude of -30?

```{r}
# I would assume that above a latitude of -30 is where x > -30
clean_data %>% 
  filter(LAT > -30) %>% 
  group_by(species, scientific_name, abbreviation) %>% 
  summarise(COUNT = sum(COUNT)) %>% 
  arrange(desc(COUNT))
```

The bird with the highest total count above a latitude of -30 is the wedge-taile shearwater (puffinus pacificus, pufpac) - with a count of 855.

## Question Four

How many different types of birds were only ever seen in groups of 1?

```{r}
# Find the average count by species
average_bird_count <- clean_data %>% 
  drop_na(COUNT) %>%
  group_by(species, scientific_name, abbreviation) %>% 
  summarise_at(vars(COUNT), list(average_count = mean)) %>% 
  filter(average_count ==  1)
```

86 different species of birds were seen on their own.

## Question Five

How many penguins were seen?
(Hint: there are many types of penguin)

```{r}
# Find the penguin species
penguins <- clean_data [grep("enguin", clean_data$species), ]
```

```{r}
# Count the penguins!
penguins  %>% 
  group_by(species) %>% 
  summarise(COUNT = sum(COUNT)) %>% 
  arrange(desc(COUNT))
```

+--------------------------------+-------------+
| Penguin Species                | Count       |
+================================+:===========:+
| Adelie penguin                 | 22          |
+--------------------------------+-------------+
| Adelie penguin SUBAD           | 5           |
+--------------------------------+-------------+
| Crested penguin (unidentified) | 1           |
+--------------------------------+-------------+
| Emperor penguin                | 10          |
+--------------------------------+-------------+
| Emperor penguin AD             | 8           |
+--------------------------------+-------------+
| Fiordland crested penguin      | 15          |
+--------------------------------+-------------+
| Little penguin                 | 90          |
+--------------------------------+-------------+
| Yellow-eyed penguin            | 3           |
+--------------------------------+-------------+
|                                |             |
+--------------------------------+-------------+
| TOTAL                          | 154         |
+--------------------------------+-------------+

: Penguin Count

In total there were 154 penguins seen.
