---
title: "R Notebook"
output: html_notebook
editor_options: 
  markdown: 
    wrap: sentence
---

# Analysis questions

For each question where we ask for bird names, give the bird's common name, scientific name and species abbreviation.

## Question One

Which bird had the most individual sightings?

```{r}
Birds_by_species <- bird_data %>% 
  group_by(Species, `Scientifc Name`, Abbreviation) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))
```

The bird with the most sightings is the Wandering albatross sensu lato PL2 (Diomedea antipodensis / exulans PL2, DIOANTEXU PL2) There were 3104 sightings.

## Question Two

Which bird had the highest total count?

```{r}
bird_data %>% 
  group_by(Species, `Scientifc Name`, Abbreviation) %>% 
  summarise(COUNT = sum(COUNT)) %>% 
  arrange(desc(COUNT))
```

The bird with the highest total count is the black-billed gull (larus bulleri, larbul) with 7732 birds counted

## Question Three

Which bird had the highest total count above a latitude of -30?

```{r}
# I would assume that above a latitude of -30 is where x > -30
full_data %>% 
  filter(LAT > -30) %>% 
  group_by(Species, `Scientifc Name`, Abbreviation) %>% 
  summarise(COUNT = sum(COUNT)) %>% 
  arrange(desc(COUNT))
```

The bird with the highest total count above a latitude of -30 is the wedge-taile shearwater (puffinus pacificus, pufpac) - with a count of 855.

## Question Four

How many different types of birds were only ever seen in groups of 1?

```{r}
# Filter by birds observed on their own
full_data %>% 
  filter(COUNT == 1) %>% 
  group_by(Species, `Scientifc Name`, Abbreviation) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))
```

298 different types of birds were seen on their own.

## Question Five

How many penguins were seen?
(Hint: there are many types of penguin)

```{r}
# Find the penguin species
penguins <- full_data [grep("peng", full_data$Species), ]
```

```{r}
# Count the penguins!
penguins <- drop_na(penguins,COUNT) %>% 
  group_by(Species) %>% 
  summarise(COUNT = sum(COUNT)) %>% 
  arrange(desc(COUNT))
```

+--------------------------------+-----------+
| Penguin Species                | Count     |
+================================+:=========:+
| Adelie penguin                 | 22        |
+--------------------------------+-----------+
| Adelie penguin SUBAD           | 5         |
+--------------------------------+-----------+
| Crested penguin (unidentified) | 1         |
+--------------------------------+-----------+
| Emperor penguin                | 10        |
+--------------------------------+-----------+
| Emperor penguin AD             | 8         |
+--------------------------------+-----------+
| Fiordland crested penguin      | 15        |
+--------------------------------+-----------+
| Little penguin                 | 90        |
+--------------------------------+-----------+
| Yellow-eyed penguin            | 3         |
+--------------------------------+-----------+
|                                |           |
+--------------------------------+-----------+
| TOTAL                          | 154       |
+--------------------------------+-----------+

: Penguin Count

In total there were 154 penguins seen.
