---
title: "Task 4 - Hallowe'en Candy Data"
output:
  html_document: 
    number_sections: yes
    toc: yes
    df_print: paged
    toc_float: yes
editor_options: 
  markdown: 
    wrap: 80
---

```{r}
library (tidyverse)
library (readxl)
library (janitor)
```

# Import Data

```{r}
candy_data <- read_csv(here("clean_data/candy_data.csv"))
```

# Question One

What is the total number of candy ratings given across the three years. 
(number of candy ratings, not number of raters. Don't count missing values)

```{r}
# For each year how many candy bars were rated in total
# Do not include missing values
candy_pivot <- candy_data %>%
  pivot_longer(cols = "100_grand_bar":"peanut_butter_jars",
            names_to = "candy",
            values_to = "rating",
            values_drop_na = TRUE
            )
```

```{r}
rating_count <- candy_pivot %>%
  group_by(year, rating) %>% 
  summarise(
    n = n()
  )  
```

## Answer:

I realised I had gone into too much detail and found out the number of ratings
by year and by response.

| Despair |  Joy   |  Meh  |   Total    |
|:-------:|:------:|:-----:|:----------:|
| 270661  | 197849 |   0   |   468510   |
|  39001  | 41889  | 37400 |   118290   |
|  58254  | 64206  | 53095 |   175555   |
| 367916  | 303944 | 90495 | **762355** |

: Number of ratings by year and response

In answer to the original question there were 762,355 responses (not including
the NAs).

# Question Two

What was the average age of people who are going out trick or treating and the
average age of people not going trick or treating?

```{r}
# Create a table for age analysis
# Drop the rows with na values in the age column from the table
age_table <- candy_data 
age_table <- drop_na(age_table,age) %>% 
  filter(age > 0)
```

```{r}
age_table %>%
  group_by(going_out) %>%
  summarise(average_age = mean(age)) 
```

| Are you going out for Hallowe'en? | Average Age |
|-----------------------------------|-------------|
| No                                | 39.13550    |
| Yes                               | 35.01952    |
| na                                | 42.30120    |

: Average age of people going trick or treating (or staying home)

# Question Three

For each of joy, despair and meh, which candy bar received the most of these
ratings?

```{r}
# Create a new table grouping the candies and counting the number of each rating
# Drop the group by to recreate it as a searchable table
candy_count <- candy_pivot %>%
  group_by(candy, rating) %>% 
  summarise(
    n = n(),.groups = "drop"
  )  
```

```{r}
# Make the table wider again to show the numbers for each rating type
candy_ratings <- candy_count %>% 
  pivot_wider(names_from = rating, values_from = n)
```

```{r}
# Find the candy bar that has received the most despair ratings
candy_ratings %>% 
  slice_max(DESPAIR)  
```

```{r}
# Find the candy bar that has received the most joy ratings
candy_ratings %>% 
  slice_max(JOY)  
```

```{r}
# Find the candy bar that has received the most meh ratings
candy_ratings %>% 
  slice_max(MEH)
```

The candy bar that has received the most 'despair' ratings is the broken glow
stick with 7905 ratings.

The candy bar that has received the most 'joy' ratings is any full sized candy
bar with 7589 ratings.

The candy bar that has received the most 'meh' ratings is lollipops with 1570
ratings.

# Question Four

How many people rated Starburst as despair?

```{r}
candy_ratings %>% 
  filter(candy == "starburst") %>% 
  select("DESPAIR")
  
```

There were 1990 'despair' rating for Starburst.

# Question Five

For the next three questions, count despair as -1, joy as +1 and meh as 0.

```{r}
# Create a new column to show the rating values
candy_pivot <- candy_pivot %>% 
  mutate(rating_value = case_when(
     (rating == "DESPAIR") ~ -1,
     (rating == "JOY") ~ 1,
     (rating == "MEH") ~ 0
  ))
```

# Question Six

What was the most popular candy bar by this rating system for each gender in the
dataset?

```{r}
candy_pivot %>% 
# Top candy bar for males
  filter(gender == "Male") %>% 
  group_by(candy) %>% 
  summarise(rating_total = sum(rating_value)) %>% 
  arrange(desc(rating_total))
```

```{r}
candy_pivot %>% 
# Top candy bar for females
  filter(gender == "Female") %>% 
  group_by(candy) %>% 
  summarise(rating_total = sum(rating_value)) %>% 
  arrange(desc(rating_total))
```

The top candy bar for both males and females is any full sized candy bar

# Question Seven

What was the most popular candy bar in each year?

```{r}
candy_pivot %>% 
# Top candy bar in 2015
  filter(year == 2015) %>% 
  group_by(candy) %>% 
  summarise(rating_total = sum(rating_value)) %>% 
  arrange(desc(rating_total))
```

```{r}
candy_pivot %>% 
# Top candy bar in 2016
  filter(year == 2016) %>% 
  group_by(candy) %>% 
  summarise(rating_total = sum(rating_value)) %>% 
  arrange(desc(rating_total))
```

```{r}
candy_pivot %>% 
# Top candy bar in 2017
  filter(year == 2017) %>% 
  group_by(candy) %>% 
  summarise(rating_total = sum(rating_value)) %>% 
  arrange(desc(rating_total))
```

Well surprise, surprise - the most popular candy bar in 2015, 2016 and 2017 is
any full sized candy bar

# Question Eight

What was the most popular candy bar by this rating for people in US, Canada, UK
and all other countries?

```{r}
candy_pivot %>% 
# Top candy bar by country
  filter(country == "usa") %>% 
  group_by(candy) %>% 
  summarise(rating_total = sum(rating_value)) %>% 
  arrange(desc(rating_total))
```

```{r}
candy_pivot %>% 
# Top candy bar by country
  filter(country == "canada") %>% 
  group_by(candy) %>% 
  summarise(rating_total = sum(rating_value)) %>% 
  arrange(desc(rating_total))
```

```{r}
candy_pivot %>% 
# Top candy bar by country
  filter(country == "uk") %>% 
  group_by(candy) %>% 
  summarise(rating_total = sum(rating_value)) %>% 
  arrange(desc(rating_total))
```

```{r}
candy_pivot %>% 
# Top candy bar by country
  filter(country == "other") %>% 
  group_by(candy) %>% 
  summarise(rating_total = sum(rating_value)) %>% 
  arrange(desc(rating_total))
```

So, again, the large candy bars are the favourite - except in the UK where hard case seems to be the preferred treat - not very tasty...