---
title: "Task 4 - Hallowe'en Candy Data"
output:
  html_document: 
    number_sections: yes
    toc: yes
    df_print: paged
    toc_float: yes
editor_options: 
  markdown: 
    wrap: 80
---

```{r}
library (tidyverse)
library (readxl)
library (janitor)
```

# Question One

What is the total number of candy ratings given across the three years. (number
of candy ratings, not number of raters. Don't count missing values)

```{r}
# For each year how many candy bars were rated in total
# Do not include missing values
candy_pivot <- combined %>%
  pivot_longer(cols = "100_grand_bar":"peanut_butter_jars",
            names_to = "code",
            values_to = "rating",
            values_drop_na = TRUE
            )
```

```{r}
rating_count <- candy_pivot %>%
  group_by(year, rating) %>% 
  summarise(
    n = n()
  )  
```

## Answer:

I realised I had gone into too much detail and found out the number of ratings
by year and by response.

| Despair |  Joy   |  Meh  |   Total    |
|:-------:|:------:|:-----:|:----------:|
| 270661  | 197849 |   0   |   468510   |
|  39001  | 41889  | 37400 |   118290   |
|  58254  | 64206  | 53095 |   175555   |
| 367916  | 303944 | 90495 | **762355** |

: Number of ratings by year and response

In answer to the original question there were 762,355 responses (not including
the NAs).

# Question Two

What was the average age of people who are going out trick or treating and the
average age of people not going trick or treating?

```{r}
combined %>%
  group_by(going_out) %>%
  summarise(average_age = mean(age)) %>% 
  arrange(desc(average_age))
```




# Question Three

For each of joy, despair and meh, which candy bar received the most of these
ratings?

# Question Four

How many people rated Starburst as despair?

# Question Five

For the next three questions, count despair as -1, joy as +1 and meh as 0.

# Question Six

What was the most popular candy bar by this rating system for each gender in the
dataset?

# Question Seven

What was the most popular candy bar in each year?

# Question Eight

What was the most popular candy bar by this rating for people in US, Canada, UK
and all other countries?
